---
title: "Longitudinal analysis of post-natal brain growth"
author: "Gareth Ball"
date: "08/05/2019"
output:   
  html_document:
    keep_md: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::read_chunk("run_models.R")
```

# Introduction

Example `R` code to perform longitudinal modelling of brain volumes as in [Ball & Seal 2019](https://link.springer.com/article/10.1007%2Fs00429-019-01829-5)

Please see the paper for details on brain tissue volume extraction and statistical modelling. Volumetric data were derived from the UNC-Wisconsin Rhesus Macaque Neurodevelopment Database. Please cite: 
>Young JT, Shi Y, Niethammer M, Grauer M, Coe CL, Lubach GR, Davis B, Budin F, Knickmeyer RC, Alexander AL, Styner MA. 2017. The UNC-Wisconsin Rhesus Macaque Neurodevelopment Database: A Structural MRI and DTI Database of Early Postnatal Development. Front Neurosci. 11.

This repository contains the code in the file `run_models.R` and the data in `brain_tissue_volumes.csv`. This
R markdown document contains a walkthrough of the procedures, including display of the code and results. It can
be reproduced by cloning the repository, loading the `Readme.Rmd` file in rstudio, and pressing the `Knit` button.

## Setup

Load required libraries. Attempt to install missing ones. Set up plotting tools.
```{r Setup, message=FALSE}

```

## Load and format data
Formatting the data units, globally setting the k for GAMS.
```{r LoadData}

```

## Intracranial volume trajectories

Test some multi-level models with combinations of age and sex, and a GAM model with a smooth function of age.
```{r FitICV}

```

Now compare the models.
```{r CompareICV}

```

Also compare to model with random slope.
```{r ICVRandomSlope}

```

Set up data for plotting
```{r ICVPlotData}

```

Perform predictions
```{r ICVPredict}

```

Create individual plots, to be combined with others below.
```{r ICVPlotsIndividual}

```

Set up plots of average trajectories, and male and female trajectories.
```{r ICVPlotsAverage}

```

```{r ICVPlotsSex}

```

Display the plots
```{r DisplayICVPlots, echo=FALSE, fig.cap="Brain volume trajectories for individuals, average, and male/female"}

```

## Grey-matter volume trajectories

Apply the same procedure to GM volumes.

First fit the models, compare, and check for random slopes.

```{r GMModels}

```

Prepare data for and predict the average trajectories.
```{r GMAverageData}

```

```{r GMAveragePredict}

```

Prepare data and predict trajectories based on sex.

```{r GMSex}

```
```{r GMSexPredict}

```
Display the trajectories for GM.

```{r GMPlotSex, echo=FALSE, fig.cap="Grey matter volume trajectories for average, and male/female"}

```


